---
title: "3VLP heatmap"
author: "PSS"
date: "2025-04-22"
output: html_document
---



**LIPID CLASS COLORS AND ORDERS**
```{r}
library(tidyverse)
library(RColorBrewer)


fontsize <- 12

#3 virus colors
colors <- c('DENV'='indianred2', 'ZIKV'='goldenrod2', 'CHIKV'='deepskyblue1', 'MOCK'='darkgray', 'NONFED'='white')
colorstreatment <- c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick4', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4')
colorsday <- c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4')

#rpColors <-  c('D7' = 'indianred1', 'D14'='indianred4', 'Z7'='goldenrod1', 'Z14'='goldenrod4', 'C7'='deepskyblue1', 'C14'='deepskyblue4', 'M7'='lightgray', 'M14' = 'darkgray')


#NS1 mouse tissue colors
# fontsize = 12
# tisColors <- c('Blood' = 'darkred', 'Brain' = 'pink', 'Small Intestine' = 'sienna3', 'Heart' = 'red', 
#                'Kidney' = 'steelblue2', 'Liver' = 'springgreen4', 'Lung' = 'orchid4', 'Spleen' = 'seagreen2' )

#classOrder <- c("TG","PC","FA", "PI", "PE", "PG", "ST", "PA", "PS", "DG", "HexCer", "CL", "Cer", "SM", "CerPE", "WE", "LPE", "LPC", "Hex2Cer", "SFE", "CAR", "LPS", "IPC", "CE", "MG", "NAE", "LPA", "SHexCer", "LPI", "FAHFA", "CoA", "FOH", "LPG", "MGDG", "CerP", "SPB","NAT", "MIPC", "SQDG", "DGCC","PIP", "SPBP", "ACer", "DGDG", "PIP2", "DGTA","DGTS", "M(IP)2C")

classFA <- c('FA', 'CAR', 'CoA', 'NAE', 'NAT')
colorsFA <- data.frame(Class = classFA, color = colorRampPalette(rev(brewer.pal(8, 'Reds')))(5))

classSP <- c('SPB', 'SPBP', 'Cer', 'ACer', 'CerP', 'IPC', 'CerPE', 'SM', 'HexCer', 'MIPC', 'M(IP)2C', 'Hex2Cer', 'SHexCer')
colorsSP <- data.frame(Class = classSP, color = colorRampPalette(rev(brewer.pal(8, 'Oranges')))(13))

classGL <- c('TG', 'DG', 'MG', 'MGDG', 'DGDG', 'SQDG', 'DGTA', 'DGTS', 'DGCC')
colorsGL <- data.frame(Class = classGL, color = colorRampPalette(rev(brewer.pal(8, 'Purples')))(9))

classGLo <- c('DG O-', 'MG O-')
colorsGLo <- data.frame(Class = classGLo, color = colorRampPalette(rev(brewer.pal(8, 'Purples')))(2))

classST <- c('ST', 'CE')
colorsST <- data.frame(Class = classST, color = c("#C51B7D" ,"#DE77AE"))

classGP <- c('PC', 'PA', 'PS', 'PE', 'PG', 'PI', 'PIP', 'CL', 'BMP', 'PEth', 'PT')
colorsGP <- data.frame(Class = classGP, color = colorRampPalette(rev(brewer.pal(8, 'Blues')))(11))

classGPo <- c('PC O-', 'PA O-', 'PS O-', 'PE O-', 'PG O-', 'PI O-')
colorsGPo <- data.frame(Class = classGPo, color = colorRampPalette(rev(brewer.pal(8, 'Blues')))(6))

classLGP <- c('LPC', 'LPA', 'LPS', 'LPE', 'LPG', 'LPI')
colorsLGP <- data.frame(Class = classLGP, color = colorRampPalette(rev(brewer.pal(8, 'Greens')))(6))

classLGPo <- c('LPC O-', 'LPA O-', 'LPS O-', 'LPE O-', 'LPG O-', 'LPI O-')
colorsLGPo <- data.frame(Class = classLGPo, color = colorRampPalette(rev(brewer.pal(8, 'Greens')))(6))

#colorsAll <- bind_rows(colorsFA, colorsGL, colorsSP, colorsST, colorsGP, colorsLGP) %>% deframe()
colorsAll <- bind_rows(colorsFA, colorsGL, colorsGLo, colorsSP, colorsST, colorsGP, colorsGPo, colorsLGP, colorsLGPo) %>% deframe()

# colorstest <- c('FA' = "#99000D", 'CAR'= "#E63228", 'CoA' = "#FB7E5E" , 'NAE'= "#FCC4AD", 'NAT'= "#FFF5F0", 'TG'= "#4A1486", 'DG'= "#66499F", 'MG'= "#7A71B4", 'MGDG'= "#928FC2",'DGDG'= "#ADABD2", 'SQGD' = "#C7C7E1", 'DGTA'= "#DFDEED", 'DGTS'= "#F0EEF6", 'DGCC'= "#FCFBFD", 'SPB'= "#8C2D04", 'SPBP'= "#B83C02",'Cer'= "#DD4D04", 'ACer'= "#EB600E", 'CerP' ="#F57520", 'IPC'= "#FC8A38", 'CerPE'= "#FD9D53", 'SM'= "#FDB06F", 'HexCer'= "#FDC48F", 'MIPC'= "#FDD5AD", 'M(IP)2C'= "#FDE2C6", 'Hex2Cer'= "#FEECDA", 'SHexCer'= "#FFF5EB", 'ST' = "#C51B7D", 'CE'="#DE77AE", 'PC'="#084594", 'PA'= "#1963AB", 'PS'= "#2E7EBB", 'PE' = "#4694C7", 'PG'= "#62A8D2", 'PI'= "#84BCDB", 'PIP'= "#A6CDE3", 'CL'="#C2D9ED", 'BMP'= "#D4E4F3", 'PEth'= "#E5EFF9", 'PT'= "#F7FBFF", 'LPC'= "#005A32", 'LPA'= "#2F974E", 'LPS'= "#69BF71", 'LPE'= "#A8DCA2", 'LPG'= "#D9F0D3", 'LPI'= "#F7FCF5")


#classOrder <- c(classFA, classGL, classSP, classST, classGP, classLGP)
classOrder <- c(classFA, classGL, classGLo, classSP, classST, classGP, classGPo, classLGP, classLGPo)

rm(colorsFA, colorsGL, colorsGLo, colorsSP, colorsST, colorsGP, colorsGPo, colorsLGP, colorsLGPo)
```


3V data
```{r}
library(tidyverse)
library(here)

dr_here() #show the project root folder for here()
here() #print the project root folder for here()

project <- '3VLP'

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
getwd()


conditionLevels <- c('NONFED0', 'MOCK0', 'MOCK5', 'MOCK7', 'MOCK14', 'MOCK21', 'DENV0', 'DENV5', 'DENV7', 'DENV14', 'DENV21', 'ZIKV0', 'ZIKV5', 'ZIKV7', 'ZIKV14', 'ZIKV21', 'CHIKV0', 'CHIKV5', 'CHIKV7', 'CHIKV14', 'CHIKV21')
conditionLevels2 <- c('NONFED0', 'MOCK0', 'DENV0', 'ZIKV0', 'CHIKV0',  'MOCK5','DENV5','ZIKV5','CHIKV5', 'MOCK7', 'DENV7', 'ZIKV7','CHIKV7', 'MOCK14',  'DENV14', 'ZIKV14',   'CHIKV14','MOCK21',  'DENV21', 'ZIKV21',  'CHIKV21')
                     

######################## Titer results from qRTPCR #########################
titer <- read_csv(here('metadata/241203 3 virus mosquito titers format.csv')) %>% 
  select(-RTPCR) %>%
  dplyr::rename(VirusTiter = Copy.Number) %>%
  mutate(DPI = factor(DPI, levels = c('5', '7', '14', '21'))) %>%
  mutate(Mosquito = paste0(Virus, DPI, '_', Sample.Number)) %>%
  select(Mosquito, VirusTiter)

meta <- read_csv(here(paste0('metadata/',project,' metadata.csv'))) %>%
    mutate(Batch = str_sub(filename, start = 20, end = 23), .after='Type') %>%
    mutate(Dilution = case_when(str_detect(filename, 'Dil1x') ~ 1,
                              str_detect(filename, 'Dil2x') ~ 0.5,
                              str_detect(filename, 'Dil4x') ~ 0.25,
                              str_detect(filename, 'Dil10x') ~ 0.1)) 

#group filtered and 2 step imputed plist with areas... NOT LOG2 transformed
plist <- read_csv(here(paste0('0 Preprocess and Analyze with GroupFiltering and 2StepImpute/250617 ',project,' data preprocessing/plist-grpfilt-impute2step-areas-',project,'.csv')))

#combine plist, metadata and titer
data <- plist %>% 
  pivot_longer(contains(':area'), names_to = 'filename', values_to = 'Abundance') %>%
    left_join(meta, by = 'filename') %>%
  left_join(titer, by = 'Mosquito') %>%
  mutate(Abundance = log2(Abundance)) %>%  #log2 transform
  mutate(Condition = factor(Condition, levels = conditionLevels))


#Lipidmaps tentative IDs
LMID <- read_csv(here(paste0('metadata/',project,' LMID.csv')))

data_LM <- data %>%
  left_join(LMID, by = 'mz') %>%
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()

LMlowdelta <- c(LMlowdelta$LMnameFeat)



#Lipidannotator MSMS IDs
LAmatch <- read_csv(here(paste0('metadata/lipidannotator-id-matched-',project,'.csv'))) %>%
  mutate(LA.Class = gsub(" .*$", "", `LA.Compound Name`)) %>%
  mutate(LA.Class = ifelse(str_detect(LA.Class, 'ACar'), 'CAR', LA.Class)) %>%
  mutate(LA.CmpdFeat = paste0(`LA.Compound Name`, ' ', feature))

data_LA <- data %>% 
left_join(LAmatch, by = 'feature')

#DPI and plot settings
whatDPI <- '_5-21_'
plotheight <- 20
#definegaps <-  c(10, 60, 110, 160) #for all DPI
# definegaps <-  c(10,20,30,40)      #for 0DPI only
 definegaps <-  c(40, 80, 120)       #for 5-21 DPI
```

Lipidannotator MSMS ID heatmap
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))


hmdatLM <- data_LA %>% filter(DPI != '0') %>%
  filter(!is.na(`LA.Compound Name`)) %>%
  select(feature, Abundance, Condition, Mosquito, Type, contains('LA.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  # mutate(FAchain = case_when(str_detect(LM.Name, 'PC O-') ~ as.numeric(str_match(LM.Name, '.*-(.+):.*')[,2]),
  #                            .default = as.numeric(str_match(LM.Name, '.* (.+):.*')[,2])),
  #        FAunsat = str_match(LM.Name, '.*:(.).*')[,2],
  #        PEorder = case_when(str_detect(LM.Name, 'PC O-') ~ 2, .default = 1)) %>%
  # arrange(LM.Class,PEorder,FAchain,FAunsat) %>%
  arrange(`LA.Compound Name`) %>%
  arrange(Condition) %>%
  distinct(LA.CmpdFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LA.CmpdFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()


plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = TRUE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_LAids_', project, whatDPI, 'clusterCols','.png'), width = 40, height = plotheight)

```



PC!
pheatmap heatmaps of raw abundance with LM tentative IDs
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()
LMlowdelta <- c(LMlowdelta$LMnameFeat)

# LMlowdelta_QCfilt <- LMlowdelta
# save(LMlowdelta_QCfilt, file = 'LMlowdelta_QCfilt_3VLP.rda')

#class
currClass <- 'PC|LPC'

#checking TG ion types and names, etc
check <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  filter(str_detect(LM.Class, currClass)) %>%
  distinct(feature, .keep_all = TRUE) %>% count(LM.Name, LM.Ion)

hmdatLM <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  filter(!str_detect(LM.Class, 'IPC'), str_detect(LM.Class, currClass), LM.Ion == '[M+H]+') %>%
  mutate(FAchain = case_when(str_detect(LM.Name, 'PC O-') ~ as.numeric(str_match(LM.Name, '.*-(.+):.*')[,2]),
                             .default = as.numeric(str_match(LM.Name, '.* (.+):.*')[,2])),
         FAunsat = str_match(LM.Name, '.*:(.).*')[,2],
         PEorder = case_when(str_detect(LM.Name, 'PC O-') ~ 2, .default = 1)) %>%
  arrange(LM.Class,PEorder,FAchain,FAunsat) %>%
  arrange(LM.Name) %>%
  arrange(Condition) %>%
  distinct(LMnameFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LMnameFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()


plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = FALSE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_', project, whatDPI, 'PC-LPC','.png'), width = 28, height = plotheight)

```

TG updated
pheatmap heatmaps of raw abundance with LM tentative IDs
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()
LMlowdelta <- c(LMlowdelta$LMnameFeat)

#class
currClass <- 'TG'

#checking TG ion types and names, etc
check <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  filter(str_detect(LM.Class, currClass)) %>%
  distinct(feature, .keep_all = TRUE) %>% count(LM.Name, LM.Ion)

hmdatLM <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  filter(str_detect(LM.Class, currClass), LM.Ion == '[M+NH4]+') %>%
  mutate(FAchain = case_when(str_detect(LM.Name, 'PC O-') ~ as.numeric(str_match(LM.Name, '.*-(.+):.*')[,2]),
                             .default = as.numeric(str_match(LM.Name, '.* (.+):.*')[,2])),
         FAunsat = str_match(LM.Name, '.*:(.).*')[,2],
         PEorder = case_when(str_detect(LM.Name, 'PC O-') ~ 2, .default = 1)) %>%
  arrange(LM.Class,PEorder,FAchain,FAunsat) %>%
  arrange(LM.Name) %>%
  arrange(Condition) %>%
  distinct(LMnameFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LMnameFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()


plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = FALSE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_', project,whatDPI, currClass, '.png'), width = 28, height = plotheight)

```



DG
pheatmap heatmaps of raw abundance with LM tentative IDs
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()
LMlowdelta <- c(LMlowdelta$LMnameFeat)

currClass <- 'DG'

#checking ion types and names, etc
check <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  filter(str_detect(LM.Class, currClass)) %>%
  distinct(feature, .keep_all = TRUE) %>% count(LM.Name, LM.Ion)


hmdatLM <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  filter(str_detect(LM.Class, currClass)) %>%
  mutate(FAchain = as.numeric(str_match(LM.Name, '.* (.+):.*')[,2]),
         FAunsat = str_match(LM.Name, '.*:(.).*')[,2],
         DGorder = case_when(str_detect(LM.Name, 'DG O-') ~ 2, .default = 1)) %>%
  arrange(LM.Class,DGorder,FAchain,FAunsat) %>%
  arrange(Condition) %>%
  distinct(LMnameFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LMnameFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()


plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = FALSE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_', project,whatDPI, currClass, '.png'), width = 28, height = plotheight)

```

FA update
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()
LMlowdelta <- c(LMlowdelta$LMnameFeat)

currClass <- 'FA'

#checking ion types and names, etc
check <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  filter(str_detect(LM.Class, currClass)) %>%
  distinct(feature, .keep_all = TRUE) %>% count(LM.Name, LM.Ion)


hmdatLM <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  filter(str_detect(LM.Class, currClass), !str_detect(LM.Name, 'FA 4:1')) %>%
  mutate(FAchain = as.numeric(str_match(LM.Name, '.*FA (.+):.*')[,2]),
         FAunsat = str_match(LM.Name, '.*:(.).*')) %>%
  arrange(LM.Class,FAchain,FAunsat) %>%
  arrange(Condition) %>%
  distinct(LMnameFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LMnameFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()


plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = FALSE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_', project,whatDPI, currClass, '.png'), width = 18, height = plotheight)

```


NAE update
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()
LMlowdelta <- c(LMlowdelta$LMnameFeat)

currClass <- 'NAE'

#checking ion types and names, etc
check <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  filter(str_detect(LM.Class, currClass)) %>%
  distinct(feature, .keep_all = TRUE) %>% count(LM.Name, LM.Ion)


hmdatLM <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  filter(str_detect(LM.Class, currClass), !str_detect(LM.Name, 'FA 4:1')) %>%
  mutate(FAchain = as.numeric(str_match(LM.Name, '.*[A-Z] (.+):.*')[,2]),
         FAunsat = str_match(LM.Name, '.*:(.).*')[,2]) %>%
  arrange(LM.Class,FAchain,FAunsat) %>%
  arrange(Condition) %>%
  distinct(LMnameFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LMnameFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()


plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = FALSE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_', project,whatDPI, currClass, '.png'), width = 8, height = plotheight)

```


CAR update
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()
LMlowdelta <- c(LMlowdelta$LMnameFeat)

currClass <- 'CAR'

#checking ion types and names, etc
check <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  filter(str_detect(LM.Class, currClass)) %>%
  distinct(feature, .keep_all = TRUE) %>% count(LM.Name, LM.Ion)


hmdatLM <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  filter(str_detect(LM.Class, currClass), !str_detect(LM.Name, 'FA 4:1')) %>%
  mutate(FAchain = as.numeric(str_match(LM.Name, '.*[A-Z] (.+):.*')[,2]),
         FAunsat = str_match(LM.Name, '.*:(.).*')[,2]) %>%
  arrange(LM.Class,FAchain,FAunsat) %>%
  arrange(Condition) %>%
  distinct(LMnameFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LMnameFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()


plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = FALSE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_', project,whatDPI, currClass, '.png'), width = 8, height = plotheight)

```



SPB and SPBP
pheatmap heatmaps of raw abundance with LM tentative IDs
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()
LMlowdelta <- c(LMlowdelta$LMnameFeat)


#class
currClass <- 'SPB|SPBP'

#checking ion types and names, etc
check <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  filter(str_detect(LM.Class, currClass)) %>%
  distinct(feature, .keep_all = TRUE) %>% count(LM.Name, LM.Ion)

#SPB 5
#SPBP 3

#Cer 40
#ACer 8
#CerP 5


#SM 31
#CerPE 46
#IPC 17

#MIPC 6
#M(IP)2C 0

#HexCer 59
#Hex2Cer 3
#SHexCer 6



hmdatLM <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  filter(str_detect(LM.Class, currClass)) %>%
  mutate(FAchain = as.numeric(str_match(LM.Name, '.*[A-Z] (.+):.*')[,2]),
         FAunsat = str_match(LM.Name, '.*:(.).*')[,2]) %>%
  arrange(FAchain,FAunsat) %>%
  arrange(Condition) %>%
  distinct(LMnameFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LMnameFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()


plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = FALSE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_', project,whatDPI, 'SPB-SPBP', '.png'), width = 5, height = plotheight)

```



Cer, ACer and CerP
pheatmap heatmaps of raw abundance with LM tentative IDs
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()
LMlowdelta <- c(LMlowdelta$LMnameFeat)


#class
currClass <- 'Cer\\b|ACer\\b|CerP\\b'

#checking TG ion types and names, etc
check <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  filter(str_detect(LM.Class, currClass), !str_detect(LM.Class, 'HexCer|SHexCer|Hex2Cer')) %>%
  distinct(feature, .keep_all = TRUE) %>% count(LM.Name, LM.Ion)

#SPB 5
#SPBP 3

#Cer 40
#ACer 8
#CerP 5


#SM 31
#CerPE 46
#IPC 17

#MIPC 6
#M(IP)2C 0

#HexCer 59
#Hex2Cer 3
#SHexCer 6



hmdatLM <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  filter(str_detect(LM.Class, currClass),!str_detect(LM.Class, 'HexCer|SHexCer|Hex2Cer')) %>%
  mutate(FAchain = as.numeric(str_match(LM.Name, '.* (.+):.*')[,2]),
         FAunsat = str_match(LM.Name, '.*:(.).*')[,2]) %>%
  arrange(FAchain,FAunsat) %>%
  arrange(Condition) %>%
  distinct(LMnameFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LMnameFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()


plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = FALSE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_', project,whatDPI, 'Cer-CerP_ACer', '.png'),  width = 16, height = plotheight)

```


SM, CerPE, IPC, MIPC
pheatmap heatmaps of raw abundance with LM tentative IDs
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()
LMlowdelta <- c(LMlowdelta$LMnameFeat)


#class
currClass <- 'SM|CerPE|IPC'

#checking TG ion types and names, etc
check <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  filter(str_detect(LM.Class, currClass)) %>%
  distinct(feature, .keep_all = TRUE) %>% count(LM.Name, LM.Ion)

#SPB 5
#SPBP 3

#Cer 40
#ACer 8
#CerP 5


#SM 31
#CerPE 46
#IPC 17
#MIPC 6
#M(IP)2C 0

#HexCer 59
#Hex2Cer 3
#SHexCer 6



hmdatLM <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  filter(str_detect(LM.Class, currClass)) %>%
  mutate(FAchain = as.numeric(str_match(LM.Name, '.* (.+):.*')[,2]),
         FAunsat = str_match(LM.Name, '.*:(.).*')[,2]) %>%
  arrange(LM.Class,FAchain,FAunsat) %>%
  arrange(Condition) %>%
  distinct(LMnameFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LMnameFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()


plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = FALSE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_', project,whatDPI, 'SM-CerPE-IPC-MIPC', '.png'),  width = 16, height = plotheight)

```


HexCer, Hex2Cer, SHexCer
pheatmap heatmaps of raw abundance with LM tentative IDs
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()
LMlowdelta <- c(LMlowdelta$LMnameFeat)


#class
currClass <- 'HexCer|Hex2Cer|SHexCer'

#checking TG ion types and names, etc
check <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  filter(str_detect(LM.Class, currClass)) %>%
  distinct(feature, .keep_all = TRUE) %>% count(LM.Name, LM.Ion)

#SPB 5
#SPBP 3

#Cer 40
#ACer 8
#CerP 5


#SM 31
#CerPE 46
#IPC 17
#MIPC 6
#M(IP)2C 0

#HexCer 59
#Hex2Cer 3
#SHexCer 6



hmdatLM <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  filter(str_detect(LM.Class, currClass)) %>%
  mutate(FAchain = as.numeric(str_match(LM.Name, '.* (.+):.*')[,2]),
         FAunsat = str_match(LM.Name, '.*:(.).*')[,2]) %>%
  arrange(LM.Class,FAchain,FAunsat) %>%
  arrange(Condition) %>%
  distinct(LMnameFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LMnameFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()



plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = FALSE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_', project,whatDPI, 'HexCer-Hex2Cer-SHexCer', '.png'),  width = 14, height = plotheight)

```


PE, LPE
pheatmap heatmaps of raw abundance with LM tentative IDs
```{r}
library(pheatmap)
colorshm <- list(DPI=c('0' = '#264653','5'='#2a9d8f','7'='#e9c46a','14'='#e76f51','21'='firebrick4'),
           Treatment = c('NONFED' = 'black', 'MOCK' = 'gray', 'DENV' = 'firebrick3', 'ZIKV' = 'darkgoldenrod4', 'CHIKV' = 'deepskyblue4'))

LMlowdelta <- data_LM %>% filter(!is.na(LM.Name)) %>%
   mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
   group_by(feature) %>% slice(which.min(LM.Delta)) %>% ungroup()
LMlowdelta <- c(LMlowdelta$LMnameFeat)


#class
currClass <- '\\bPE|LPE'

#checking TG ion types and names, etc
check <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  filter(str_detect(LM.Class, currClass)) %>%
  distinct(feature, .keep_all = TRUE) %>% count(LM.Name, LM.Ion)





hmdatLM <- data_LM %>% 
  mutate(LMnameFeat = paste0(LM.Name, ' ', LM.Ion, ' ', feature)) %>%
  filter(LMnameFeat %in% LMlowdelta) %>%
  select(feature, LMnameFeat, Abundance, Condition, Mosquito, Type, contains('LM.'), filename) %>%
  mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
  filter(Type == 'Exp') %>% 
  filter(str_detect(LM.Class, currClass)) %>%
  mutate(FAchain = case_when(str_detect(LM.Name, 'PE O-') ~ as.numeric(str_match(LM.Name, '.*-(.+):.*')[,2]),
                             .default = as.numeric(str_match(LM.Name, '.* (.+):.*')[,2])),
         FAunsat = str_match(LM.Name, '.*:(.).*')[,2],
         PEorder = case_when(str_detect(LM.Name, 'PE O-') ~ 2, .default = 1)) %>%
  arrange(LM.Class,PEorder,FAchain,FAunsat) %>%
  arrange(Condition) %>%
  distinct(LMnameFeat, Mosquito, Abundance) %>%
  pivot_wider(names_from = LMnameFeat, values_from = Abundance) %>%
  column_to_rownames('Mosquito')

hmdatLM <- as.matrix(hmdatLM)

groups <- data_LM %>% select(Mosquito, Condition, DPI, Treatment) %>% distinct(Mosquito, .keep_all = TRUE) %>% filter(!is.na(Mosquito)) %>%
 mutate(Condition = factor(Condition, levels = conditionLevels)) %>%
 arrange(Condition) %>% 
 select(Mosquito, DPI, Treatment) %>%
column_to_rownames('Mosquito') %>% as.data.frame()



plot1 <- pheatmap(hmdatLM, scale = 'column', cluster_rows = FALSE, cluster_cols = FALSE,
                  gaps_row = definegaps,
                  annotation_row = groups,
                  annotation_colors = colorshm,
                  color = colorRampPalette(c('blue','white', 'red'))(50))


ggsave(plot = plot1, paste0('hm_', project,whatDPI, 'PE-LPE', '.png'),  width = 14, height = plotheight)

```



