---
title: "3VLP preprocessing"
author: "PSS"
date: "2025-06-17"
output: word_document
---

The plan in this rmd is to perform FOR A GIVEN DATASET (project); e.g. 3VLP, 3VLN, 3VMP, 3VMN

CHUNK1:
>call in original (unchanged) mzmine peaklist output
>generate metadata; filename and its related to Type, Mosquito, Condition, DPI, Treatment, Group

CHUNK2:
>group based presence filtering
>2 step imputation (1/5 min and knn based on group presence)
>generate two peaklists from the group filtered and 2step imputed
   "-areas" that will contain feature identifiers and peak areas
   '-full' that will contain feature identifiers and all info that pops out of mzmine

CHUNK 3:
>limma analysis

CHUNK 4:
>limma results: 
  define significant up/down differential abundance
  concatenate results into fewer .csv files

CHUNK 5:
>lipidannotator results

CHUNK 6:
>LIPIDMAPS tentative ID results




CHUNK 1
import original mzmine output peak list
generate metadata for sample types and conditions, etc
```{r}
library(tidyverse)
library(here)

dr_here() #show the project root folder for here()
here() #print the project root folder for here()

project <- '3VLP'

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
getwd()


#original mzmine plist
plist <- read_csv(here('241121_3Virus_mzmine_lipids_pos/Aligned_feature_list_gapfill_dup_rmblank-pList_corr_quant_modular.csv')) %>%
  #this next line is specific to 3VLP; adjust/remove for other projects
  setNames(gsub('LPB2_055_ZIKV21', 'LPB2_055_ZIKV21_1', names(.))) %>%
  mutate(feature = paste0('MZ', round(mz, 4), '_', round(rt, 2)), .after = 'id') 


#create metadata from original mzmine plist
meta <- plist %>%
  select(id,feature,mz,rt, contains(':area')) %>%
  pivot_longer(contains(':area'), names_to = 'filename', values_to = 'Abundance') %>%
  relocate(c('filename', 'Abundance'), .after = 'rt') %>%
  mutate(Type = case_when(str_detect(filename, 'pQC\\b') ~ 'QC',  #use 'pQC\\b' if pQC is the end of the filename string (probably have to str_sub(filename) to get this)
                                        str_detect(filename, 'DDA') ~ 'DDA',
                                        str_detect(filename, 'Extr') ~ 'ExtrBlank',
                                        str_detect(filename, 'Conc') ~ 'QCDil',
                                          .default = 'Exp'), .after = 'filename') %>%
  mutate(Batch = str_sub(filename, start = 20, end = 23), .after='Type') %>%
  mutate(Mosquito = str_sub(filename, start = 29, end = -11), 
         Mosquito = ifelse(str_detect(Mosquito, 'pQC|ExtrBl'), NA_character_, Mosquito), .after='Type') %>%
  mutate(Condition = str_extract(Mosquito, ".+?(?=_)"),
         DPI = as.double(str_extract(Mosquito, '[0-9]+')),
         Treatment = str_extract(Mosquito, "[A-Z]+"), .after = 'Mosquito') %>%
  mutate(Group = Condition) %>%
  select(filename, Type, Mosquito, Condition, DPI, Treatment, Group) %>%
  distinct(filename, .keep_all = TRUE)

#import metadata created prior; they are identical
#metaimport <- read_csv(here('metadata/3VLP metadata.csv')) %>% mutate(Group = Condition)


```

CHUNK 2
Experimental samples only
Group filter and 2 step impute
```{r}
#impute by 1/5 min, followed by knn

#"my_data" is long form with id, sample name, and peak area

# in these steps, I'm not using the pooled QC, but I often have included them in a quality-control step
# also, I have interim objects because I use information from these for building tables, etc in reports


####### STEP 1
# remove any features that are missing in all samples
# these features are generally introduced by blanks / additional samples that don't match the treatments of interest

my_data <- plist %>% select(id, contains(':area')) %>%
  pivot_longer(contains(':area'), names_to = 'filename', values_to = 'area')

all_missing <- my_data %>% 
  group_by(id) %>%
  mutate(all_na = sum(is.na(area)) == n()) %>%
  filter(all_na == TRUE) %>% 
  count(id)

clean_data <-  my_data %>%
  filter(!(id %in% all_missing$id))


######## STEP 2
#find features that are missing in more than half of the samples in all groups

#find number of samples with missing value for each treatment group x feature id
missing_for_imputation <- clean_data %>%
  left_join(meta %>% select(filename, Mosquito, Group), by = 'filename') %>%
  filter(!is.na(Group)) %>% #filter out sample Types: Blank, QC, QCDil, DDA
  group_by(Group, id) %>%
  summarise(n_missing = sum(is.na(area)))

#pull features missing in >= 50% of the treatment group
missing_in_group <- missing_for_imputation %>%
  group_by(id) %>%
  mutate(groups_missing = sum(n_missing > 5)) %>% #set presence threshold here; 3V data has 10 samples per group so 5 is 50%
  pivot_wider(names_from = Group, values_from = n_missing) %>%
  filter(groups_missing == 21) #== 21  means feature is less than 50% present in all experimental groups

#remove features missing in all groups (21 groups for 3 virus data)
clean_data <- clean_data %>%
  filter(!id %in% missing_in_group$id)


####### STEP 3
# identify imputation method for each missing feature
missing_for_imputation <- missing_for_imputation %>%
  filter(!id %in% missing_in_group$id) %>%
  #cutoffs based on n=10 per group
  mutate(impute_method = case_when(n_missing == 0 ~ 'none',
                                   n_missing > 2 ~ 'min',
                                   n_missing <= 2 ~ 'knn'))


####### STEP 4
# add impute method to data, and impute 1/5 minimum
set.seed(272)

imputed_data <- clean_data %>%
  left_join(meta %>% select(filename, Mosquito, Group), by = 'filename') %>%
  left_join(missing_for_imputation %>% select(-n_missing), by = c('id', 'Group')) %>%
  filter(!is.na(Group)) %>%
  group_by(id) %>%
  mutate(min_area = min(area, na.rm = TRUE) / 5,
         min_area = jitter(min_area, 1),
         area = ifelse(impute_method == 'min' & is.na(area), min_area, area)) %>%
  ungroup() %>%
  select(-c(min_area))


####### STEP 5
# impute knn for the remaining missing values

impute_knn <- imputed_data %>% 
  select(-c(Mosquito, Group, impute_method)) %>%
  pivot_wider(names_from = filename, values_from = area) %>%
  as.data.frame() %>%
  column_to_rownames('id') %>%
  as.matrix() %>%
  impute::impute.knn(k = 5, rowmax = 0.2, colmax = 0.8, rng.seed = 12345)

knn_imputed <- impute_knn$data %>%
  as.data.frame() %>%
  rownames_to_column('id') %>%
  as_tibble() %>%
  mutate(id = as.double(id))
  #pivot_longer(-id, names_to = 'filename', values_to = 'area')

####### STEP 6
# create and export group filtered, 2 step imputed peaklists 

#full; with all mzmine output information
plist_filter_impute_full <- plist %>% 
  select(-contains(':area')) %>%
  filter(id %in% knn_imputed$id) %>%
  left_join(knn_imputed, by = 'id')
  
#write_csv(plist_filter_impute_full, paste0('plist-grpfilt-impute2step-full-', project, '.csv'))
  
#areas; with only feature identifiers and areas
plist_filter_impute_areas <- plist %>% 
  select(-contains(':area')) %>%
  filter(id %in% knn_imputed$id) %>%
  left_join(knn_imputed, by = 'id') %>% 
  select(id, feature, mz, rt, contains(':area'))
  
#write_csv(plist_filter_impute_areas, paste0('plist-grpfilt-impute2step-areas-', project, '.csv'))

```

CHUNK 3
limma
export limma results
export mummichog input formats
```{r}


# data for limma is wide with features in row and samples in columns
# the feature names go into rownames
# no other columns should be in the data

limma_data <- plist_filter_impute_areas %>%
  pivot_longer(contains(':area'), names_to = 'filename', values_to = 'area') %>%
  mutate(area = log2(area)) %>% #limma requires log2 values
  pivot_wider(names_from = 'filename', values_from = 'area') %>%
  select(feature, contains(':area')) %>%
  as.data.frame %>%
  column_to_rownames("feature")


# limma also requires a vector showing which group the samples are in
# this needs to be in the same order as the column names
# groups need to be factors
# if you want to combine groups, you will need to do that at this stage
# I include an example called new group

limma_groups <- meta %>% distinct(filename, .keep_all = TRUE) %>% filter(!str_detect(filename, 'QC|ExtrB')) %>%
  select(filename, Condition) %>%
  mutate(Condition = factor(Condition)) %>%
  rename(dataFileName = filename)
  

identical(limma_groups$dataFileName, names(limma_data))

### limma analysis gives pairwise fold change, t statistic and p-value
# if you don't have limma, install it from Bioconductor
# if you don't have BiocManager, you will need to install it first

# install.packages("BiocManager")
# BiocManager::install("limma")

library(limma)

################ create the design matrix

limma_design <- model.matrix(~ 0 + limma_groups$Condition)
colnames(limma_design) <- levels(limma_groups$Condition)



############# define the pairwise contrasts to include
############# limma has a maximum of 5 pairs
############# in my example I compare each of the dengue groups to non-dengue

contrasts <- makeContrasts(MOCK0 - NONFED0, DENV0 - MOCK0, DENV5 - MOCK5, DENV7 - MOCK7, DENV14 - MOCK14, DENV21 - MOCK21, ZIKV0 - MOCK0, ZIKV5 - MOCK5, ZIKV7 - MOCK7, ZIKV14 - MOCK14, ZIKV21 - MOCK21, CHIKV0 - MOCK0, CHIKV5 - MOCK5, CHIKV7 - MOCK7, CHIKV14 - MOCK14, CHIKV21 - MOCK21, levels = limma_design)

# MOCK0 - NONFED0, DENV0 - MOCK0, DENV5 - MOCK5, DENV7 - MOCK7, DENV14 - MOCK14, DENV21 - MOCK21, ZIKV0 - MOCK0, ZIKV5 - MOCK5, ZIKV7 - MOCK7, ZIKV14 - MOCK14, ZIKV21 - MOCK21, CHIKV0 - MOCK0, CHIKV5 - MOCK5, CHIKV7 - MOCK7, CHIKV14 - MOCK14, CHIKV21 - MOCK21

############ fit the initial model


fit <- lmFit(limma_data, limma_design)


############ add the contrasts and make eBayes adjustment


fit <- contrasts.fit(fit, contrasts)
fit <- eBayes(fit)

##  ##
##  ##
##  ##

#### LIMMA RESULTS
## to get the full information for fit:
# first put the rownames into its own column, or you will lose them

#Results all contrasts combined
results <- topTable(fit, n = Inf) %>%
  rownames_to_column('feature') %>%
  as_tibble 

write_csv(results, paste0('limma_', project, '_allContrasts.csv'))


#Results DENV contrasts
DEN_results <- topTable(fit, coef = c(2,3,4,5,6), n = Inf) %>%
  rownames_to_column('feature') %>%
  as_tibble

write_csv(DEN_results, paste0('limma_', project, '_DENV.csv'))

#Results ZIKV contrasts
ZIKV_results <- topTable(fit, coef = c(7,8,9,10,11), n = Inf) %>%
  rownames_to_column('feature') %>%
  as_tibble 

write_csv(ZIKV_results, paste0('limma_', project, '_ZIKV.csv'))

#Results CHIKV contrasts
CHIKV_results <- topTable(fit, coef = c(12,13,14,15,16), n = Inf) %>%
  rownames_to_column('feature') %>%
  as_tibble 

write_csv(CHIKV_results, paste0('limma_', project, '_CHIKV.csv'))

##  ##
##  ##
##  ##


###write individual contrast results to .csv

comparisons <- c('M0NF0', 'D0M0', 'D5M5', 'D7M7', 'D14M14', 'D21M21', 'Z0M0', 'Z5M5', 'Z7M7', 'Z14M14', 'Z21M21', 'C0M0', 'C5M5', 'C7M7', 'C14M14', 'C21M21')

for(cont in 1:ncol(contrasts)){

  print(colnames(contrasts)[cont])
  
  contname <- gsub(' - ', '...', colnames(contrasts)[cont]) #dont need to use this because im building a comparison name vector
  print(contname)
  
  print(paste0('COMPARISON:', comparisons[cont]))
  
  
  result <- topTable(fit, coef = cont, n = Inf) %>%
    rownames_to_column('feature') %>%
    as_tibble 
  
  write_csv(result, file = paste0('limma_',project, '_', comparisons[cont], '.csv'))
}

##  ##
##  ##
##  ##

###create mummichog inputs from limma results

comparisons <- c('M0NF0', 'D0M0', 'D5M5', 'D7M7', 'D14M14', 'D21M21', 'Z0M0', 'Z5M5', 'Z7M7', 'Z14M14', 'Z21M21', 'C0M0', 'C5M5', 'C7M7', 'C14M14', 'C21M21')

for(cont in 1:ncol(contrasts)){

  print(colnames(contrasts)[cont])
  
  contname <- gsub(' - ', '...', colnames(contrasts)[cont]) #dont need to use this because im building a comparison name vector
  print(contname)
  
  print(paste0('COMPARISON:', comparisons[cont]))
  
  
  result <- topTable(fit, coef = cont, n = Inf) %>%
    rownames_to_column('feature') %>%
    as_tibble %>%
    mutate(m.z = str_match(feature, 'MZ(.*?)_')[,2],
           r.t = sub('.*\\_', '', feature),
           p.value = adj.P.Val,
           t.score = t) %>%
    select(m.z, p.value, t.score, r.t)
    
  
  write_csv(result, file = paste0('mummichogInput_limma_', project, '_', comparisons[cont], '.csv'))
}

```


CHUNK 4
pull in limma results for PAIRWISE contrasts then concatenate results
define significant up/down differential abundance
log2(0.585) = 1.5 fold change
```{r}

comparisons <- c('M0NF0', 'D0M0', 'D5M5', 'D7M7', 'D14M14', 'D21M21', 'Z0M0', 'Z5M5', 'Z7M7', 'Z14M14', 'Z21M21', 'C0M0', 'C5M5', 'C7M7', 'C14M14', 'C21M21')

for(comp in comparisons){
  
  limma <- read_csv(paste0('limma_',project,'_', comp, '.csv')) %>%
  mutate(DIRECTION = ifelse(.[2] > 0.585, 'UP', 'DOWN')[,1],
         DIRECTION = ifelse(.[2] < -0.585, 'DOWN', DIRECTION),
         DIRECTION = ifelse(.[2] < 0.585 & .[2] > -0.585, 'NONE', DIRECTION)[,1]) %>%
  mutate(Prob = B/(1+B)) %>%
  dplyr::rename( !! paste0(comp, '_log2FC') := 2,
          !! paste0(comp, '_AveExpr') := 3,
          !! paste0(comp, '_t') := 4,
          !! paste0(comp, '_Pval') := 5,
          !! paste0(comp, '_adjPval') := 6,
          !! paste0(comp, '_B') := 7,
          !! paste0(comp, '_Dir') := 8,
          !! paste0(comp, '_Prob') := 9)
  
  assign(paste0('limma_', comp), limma)
  rm(limma)
}


pw_limma_DENV <- plyr::join_all(list(limma_D0M0,limma_D5M5,limma_D7M7,limma_D14M14,limma_D21M21), by = 'feature', type = 'left')

pw_limma_ZIKV <- plyr::join_all(list(limma_Z0M0,limma_Z5M5,limma_Z7M7,limma_Z14M14,limma_Z21M21), by = 'feature', type = 'left')

pw_limma_CHIKV <- plyr::join_all(list(limma_C0M0,limma_C5M5,limma_C7M7,limma_C14M14,limma_C21M21), by = 'feature', type = 'left')

pw_limma_NONFED <- limma_M0NF0


write_csv(pw_limma_DENV, paste0('limma results_', project, '_pairwise all DENV vs MOCK.csv'))
write_csv(pw_limma_ZIKV, paste0('limma results_', project, '_pairwise all ZIKV vs MOCK.csv'))
write_csv(pw_limma_CHIKV, paste0('limma results_', project,'_pairwise all CHIKV vs MOCK.csv'))
write_csv(pw_limma_NONFED, paste0('limma results_', project, '_pairwise MOCK vs NONFED.csv'))


```


CHUNK 5
fuzzy join lipidannotator MSMS identifications
```{r}

#lipidannotator IDs
LAid <- read_csv(here('metadata/database-search-results/250114 3V LP lipidannotator.csv')) %>% select(-c('DT', 'CCS', 'Gas')) %>%
  filter(!is.na(Score))
colnames(LAid) <- paste('LA.', colnames(LAid), sep='')

LAid <- LAid %>%
 #mutate(feature = paste0('MZ', round(LA.mz, 4), '_', round(LA.RT, 2)), .after = 'LA.Feat ID') %>%
 mutate(mz = LA.mz, rt =LA.RT, .before = `LA.Feat ID`)


#functions for fuzzy join
match_ppm <- function(x, y) {
  abs(x - y) / x * 10^6 <= 10 }

match_rt <- function(x, y) {
  abs(x - y) <= 0.1 }

#fuzzy join
fuzzy <- fuzzyjoin::fuzzy_join(x = plist_filter_impute_areas %>% select(-contains(':area')), y = LAid,
  by = c("mz", "rt"),
  match_fun = list(match_ppm, match_rt),
  mode = "full")

fuzzy <- fuzzy %>% select(-mz.y, -rt.y) %>% rename(mz = mz.x, rt = rt.x)
  
#investigate matches
LA_not_matched <- fuzzy %>% filter(is.na(feature)) %>% distinct(`LA.Feat ID`, .keep_all = TRUE)

LA_matched <- fuzzy %>% filter(!is.na(feature)) %>% distinct(`LA.Feat ID`, .keep_all = TRUE)

LA_duplicated <- fuzzy %>% filter(!is.na(feature), !is.na(`LA.Feat ID`)) %>%
  add_count(feature, name = 'n_feat') %>%
  add_count(`LA.Feat ID`, name = 'n_LAid') %>% filter(n_feat>1 | n_LAid > 1) %>%
  relocate(c('n_feat', 'n_LAid'), .after = 'feature')

 write_csv(LA_not_matched, paste0('lipidannotator-id-not-matched-',project, '.csv'))
 write_csv(LA_matched, paste0('lipidannotator-id-matched-',project, '.csv'))
 write_csv(LA_duplicated, paste0('lipidannotator-id-duplicatematches-',project, '.csv'))




```

CHUNK 6
join tentative IDs from LIPIDMAPS
```{r}

###### lipidmaps tentative ID search 3VLP all features######
LMID_1 <- read_tsv(here('metadata/database-search-results/3VLP_allfeats_LMsearch_first4999.tsv'))
LMID_2<- read_tsv(here('metadata/database-search-results/3VLP_allfeats_LMsearch_last4000.tsv'))

LMID_all <- bind_rows(LMID_1, LMID_2) %>%
  rename(InputMass = `Input Mass`, MatchedMass = `Matched Mass`, Examples = `LMSD Examples`)

colnames(LMID_all) <- paste('LM.', colnames(LMID_all), sep = '')

LMID_all <- LMID_all %>%
  mutate(mz = LM.InputMass, .before = LM.InputMass) %>%
  mutate(LM.Class = gsub(" .*$", "", LM.Name), .before = LM.Name)

rm(LMID_1, LMID_2)

write_csv(LMID_all, paste0(project,' LMID.csv'))




```























